Clustering time series based on probability distributions across temporal granularities

### 2021-11-17 main.Rmd

The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee). Half-hourly measurements of usage for more than 13,000 household electricity smart meters is provided from from October 2011 to March 2014. <!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning. The behavioral patterns differ amongst customers due to many temporal dependencies. Some households use a dryer, while others dry their clothes on a line. Their weekly usage profile may reflect this. They may vary monthly, with some customers using more air conditioners or heaters than others, while having equivalent electrical equipment and weather circumstances. Some customers are night owls, while others are morning larks. Daily energy usage varies depending on whether customers stay home or work away from home. Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior complex. The challenge is to be able to cluster consumers into these type of expected patterns, and other unexpected patterns, using only their energy usage history [@Ushakova2020-rl]. <!--To safeguard the customers' privacy, it is probable that such information is not accessible. Also, energy suppliers may not always update client information, such as property features, in a timely manner.--> There is a growing need to have methods that can examine the energy usage heterogeneity observed in smart meter data and what are some of the most common power consumption patterns.

@@

df = tibble::tibble(scenario, designs, niter, mu, nT) %>% rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
  #tibble::tibble(scenario, " designs" = designs, "R" =niter,  "$\\mu$" = mu, "T"=nT) %>% 
  knitr::kable(format = "latex",
             escape = FALSE,
             caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($R$) for each design, differences between means ($\\mu$) across granularities and series lengths ($T$).")%>%
  kableExtra::collapse_rows(columns = 3:5)

df
```

@@

Huge data sets present more complications for clustering. Clustering algorithms work well when there are well-separated clusters, with no nuisance variables or nuisance observations. When converting a series to granularities, many variables (each level of a granularity) are generated, possibly creating a slew of nuisance variables. Some households may have a mix of energy use patterns, which could be considered nuisance observations located between major clusters. For this reason, we have chosen to select a small group of households with relatively distinct and different patterns in order to illustrate the clustering more simply. This is often a good approach to tackling a big analysis task, to start with a simpler task. The approach, however, is applicable to all households.

@@

The steps for household filtering and variable selection were:

@@

## Choosing households {#sec:prototype}

>>>>

The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee). Half-hourly measurements of usage for more than 13,000 electricity smart meter customers is provided from from October 2011 to March 2014. <!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Customers vary in size, location, and amenities such as solar panels, central heating, and air conditioning. The behavioral patterns differ amongst customers due to many temporal dependencies. Some customers use a dryer, while others dry their clothes on a line. Their weekly usage profile may reflect this. They may vary monthly, with some customers using more air conditioners or heaters than others, while having equivalent electrical equipment and weather circumstances. Some customers are night owls, while others are morning larks. Daily energy usage varies depending on whether customers stay home or work away from home. Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior complex. The challenge is to be able to cluster consumers into these type of expected patterns, and other unexpected patterns, using only their energy usage history [@Ushakova2020-rl]. <!--To safeguard the customers' privacy, it is probable that such information is not accessible. Also, energy suppliers may not always update client information, such as property features, in a timely manner.--> There is a growing need to have methods that can examine the energy usage heterogeneity observed in smart meter data and what are some of the most common power consumption patterns.

@@

tibble::tibble(scenario, designs, niter, mu, nT) %>% #rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
  #tibble::tibble(scenario, " designs" = designs, "R" =niter,  "$\\mu$" = mu, "T"=nT) %>% 
  knitr::kable(format = "latex",
             escape = FALSE,
             caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($R$) for each design, differences between means ($\\mu$) across granularities and series lengths ($T$).")%>%
  kableExtra::collapse_rows(columns = 3:5)
```

@@

Huge data sets present more complications for clustering. Clustering algorithms work well when there are well-separated clusters, with no nuisance variables or nuisance observations. When converting a series to granularities, many variables (each level of a granularity) are generated, possibly creating a slew of nuisance variables. Some customers may have a mix of energy use patterns, which could be considered nuisance observations located between major clusters. For this reason, we have chosen to select a small group of customers with relatively distinct and different patterns in order to illustrate the clustering more simply. This is often a good approach to tackling a big analysis task, to start with a simpler task. The approach, however, is applicable to all customers.

@@

The steps for customer filtering and variable selection were:

@@

## Choosing customers {#sec:prototype}
